---
title: "Analysis LR field of large dataset using spWAVE"
output: 
  rmarkdown::html_document:
    theme: journal
    toc: yes
author: Kai Li
date: 2025-04-07
mainfont: arial
vignette: >
  %\VignetteIndexEntry{Analysis LR field of large dataset using spWAVE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  comment = "#>",
  root.dir = './'
)
```
In this vignette, we will focus on large dataset processing and field estimation.
Such large datasets are often single cell resolutions and have large number of cells but 
limited transcript coverage, like 10x Xenium and MERFISH/MERSCOPE.
To reduce the compute complexity and resource consumption, 
we will introduce an alternative "holed manner" method.
In short, for those out of range(e.g. >500 micrometer) cells, we aggregate them into multiple meta-cells 
and keep all in-range single cells to calculate the interaction field.

Besides, we will show some methods to create spWAVE object from different sources and
select ROI to estimate the field.



Let's start by loading the required libraries.
```{r setup, message = FALSE,eval=FALSE}
library(magrittr)
library(dplyr)
library(tidyr)
library(stringr)

library(Seurat)
library(Matrix)
library(spWAVE)
```

## 10x Xenium Breast Cancer dataset
First, we start with a 10x Xenium dataset of human breast cancer. 
The rawdata and metadata are downloaded from [10x Genomics](https://www.10xgenomics.com/products/xenium-in-situ/preview-dataset-human-breast), 
and more information can be found in [this article](https://www.nature.com/articles/s41467-023-43458-x)
The pre-process steps are followed the Seurat workflow and not shown here.
Basically, we just need the normalized expression matrix and spatial coordinates.
Therefore, only the SCTransform is needed.
```{r,eval=FALSE}
xen_brca_ft <- readRDS("./xen_hBRCA_tls_ft_anno.rds")

cropped.coords <- Crop(xen_brca_ft[["fov"]], x = c(0,4000), y = c(0, 4000))
xen_brca_ft[["crop"]] <- cropped.coords
ImageDimPlot(xen_brca_ft, fov = "crop",group.by = "Cluster")
ROI_barcode <- xen_brca_ft[["crop"]]@boundaries$centroids@cells

human_db <- load_database(db_source = "CellChat",db_species = "human")
hBRCA_set <- create_spWAVE_object(xen_brca_ft,human_db)

#generate meta cells expression and center coordinates
hBRCA_set %<>% generate_kmeans_coord(centers = ncol(xen_brca_ft)/25,iter.max=15,nstart=3)
hBRCA_set %<>% generate_meta_expr()

hBRCA_set %<>% perform_LR_field_hole_calc(
  ROI_barcode=ROI_barcode
)
hBRCA_set %<>% perform_S2S_score_calc(ROI_barcode=ROI_barcode)
hBRCA_set %<>%
  perform_C2C_score_calc(ROI_barcode=ROI_barcode,shuffle_iter=200,verbose=T)

```
Here, we showed the first approach to reduced the computing area by giving the barcode of cells in ROI.
This approach still includes **all cells** but only computes the field in the ROI, 
which means the cells out of the ROI are aggregated into meta-cells.
The benefit of this methods is that it keeps the information from cells out of ROI.
However, it should be noted that even when cells outside the ROI are aggregated into meta-cells, 
they remain involved in computations. 
For a very large dataset, this may result in huge distance and interaction matrices, 
and significantly increase computational resource consumption.

Again, we recommend diet the result to avoid taking too much memory and disk space.
The S2S_score will be huge with a large number of cells, 
since it's a pairwise interaction score for each cell.
```{r,eval=FALSE}
hBRCA_set@S2S_score <- list()
gc()
saveRDS(hBRCA_set,"path/you/like.rds")
```

Here is some tips for displaying the field for high resolution single cells ST dataset.
We recommend using the 'hexified' argument to plot hex-heatmap for values like field potential and expression.
If Seurat Object is given, we can use the 'image' argument to plot Seurat cluster,
and the background will be black.
Further, if not limiting the range of the coordinates, the arrow will only be displayed in the ROI.
So we add xlim and ylim to determine the range.
```{r hBRCA fig,fig.retina = 3,fig.asp=0.5,fig.width=14,fig.align = "center"}
library(patchwork)
xen_brca_ft <- readRDS("D:\\calc_temp\\field_test\\xen_hBRCA_tls_ft_anno.rds")
hBRCA_set <- readRDS("D:\\calc_temp\\field_test\\xen_hBRCA_tls_ft_spWAVE_c2c_only.rds")

field_cxcr4 <-
  perform_field_extract(hBRCA_set,LR="CXCL12.CXCR4")

p1 <- 
plot_field_direction2(field_cxcr4,point="KU",hexified = TRUE,
  arrow_linewidth=0.5,
  arrow_color="black",
  hex_binwidth=c(25,25)
  ) +
  scale_fill_gradient2(low="#3B50C5", mid="#DCDCDC",high="#AD0C27")+
  xlim(0,4000)+ylim(0,4000)

p2 <- 
plot_field_direction2(field_cxcr4,seurat_obj = xen_brca_ft,image = "cluster",
  grid_density = 0.5,
  arrow_linewidth=0.5,
  arrow_color = "cyan",
  shadow_color = "yellow",
  #arrow_alpha = 0.7,
  ) +
  xlim(0,4000)+ylim(0,4000)

p1|p2
```

```{r,clear env,include=FALSE}
rm(list=ls())
gc()

```


## smFISH/MERFISH/MERRSCOPE dataset
Next, we introduce a series datasets of human high grade serous carcinoma(HGSC). 
The processed Seurat object can obtained from [Zenodo](https://zenodo.org/records/12613839), 
and more information can be found in article [Yeh et al., 2024](https://www.nature.com/articles/s41590-024-01943-5)
Here, we used the smiFISH ST data of HGSC2 in 'Test2' group.
The coordinates of this dataset is special.
It stored in the metadata but not 
We have to manually extract it and create spWAVE object.
```{r preview HGSC2}
HGSC2 <- readRDS("D:\\calc_temp\\field_test\\ST_Test2_HGSC2_so.rds")
HGSC2@meta.data %>% head
HGSC2@meta.data[,c("x","y")] %>% summary()
```


As the coordinates range shown, this dataset covers a large spatial area. 
Thus we selected a ROI and filtered it.
Due to the nature of smFISH and cell segmentation algorithms, 
overlapping cells may occur in this dataset.
We remove those cells with distance less than 20 units (equivalent to 2 µm) 
Then, we normalize the expression of the remaining cells.
Finally, spatial coordinates and cell annotation are extracted and converted to standardized units.

In the next code chunk, we will show detail code for this process. Each steps will be commented.
```{r ROI pre-process, eval=FALSE}

#Select ROI and filter
ROI_barcode <-
HGSC2@meta.data %>%
  filter(x>23000,x<45000,y>-4300,y<10500) %>%
  rownames()

ROI <- subset(HGSC2, cells=ROI_barcode)

#ROI %>% VlnPlot(features = c("nFeature_RNA", "nCount_RNA"), ncol = 2, pt.size = 0)

ROI <- subset(ROI, subset = nCount_RNA < 3000)
ROI <- subset(ROI, subset = nFeature_RNA < 400)

#filter out overlap cells
df_ranged <- ROI@meta.data
coords <- cbind(df_ranged$x, df_ranged$y)
dmat <- as.matrix(dist(coords))
diag(dmat) <- NA
too_close <- which(apply(dmat, 1, min, na.rm=TRUE) < 20)
df_filtered <- df_ranged[-too_close, ]
ROI <- subset(ROI,cells=rownames(df_filtered))

#Normalize expr mat
ROI %<>% SCTransform()

#rescale coords to micrometer
ROI_coord  <-
  ROI@meta.data[,c("x","y")] %>%
  mutate(y=y-min(y)) %>%
  mutate(x=x/10,y=y/10)
ROI_coord$barcode <- rownames(ROI_coord)

#merge cell subtypes annotation
ROI@meta.data %<>%
  mutate(subtype=gsub("_LC","",cell.subtypes))

```

After pre-processing, we are ready to create spWAVE object and calculate the LR field.
Here, we create by manually setting the parameters of coord and cluster.
```{r HGSC2 spWAVE, eval=FALSE}
human_db <- load_database(db_source = "CellChat",db_species = "human")

ROI_spWAVE <- create_spWAVE_object(
  seurat_obj = ROI,
  database = human_db,
  coord = ROI_coord,
  cluster="subtype"
  )

ROI_spWAVE %<>% generate_kmeans_coord(centers=ncol(ROI)/20,iter.max=15,nstart=3)
ROI_spWAVE %<>% generate_meta_expr()

library(future)
future::plan("multicore",workers=16)
options(future.globals.maxSize = 100*1024^3)
ROI_spWAVE %<>% perform_LR_field_hole_calc()
ROI_spWAVE %<>% perform_S2S_score_calc()
ROI_spWAVE %<>% perform_C2C_score_calc(shuffle_iter=200,verbose=T)

```

Now, we can visualize the field result.
Here we show the projection of "CXCL12-CXCR4" and "CXCL16-CXCR6" and compare them.
```{r field proj, fig.retina = 3,fig.asp=0.6,fig.width=14,fig.align = "center"}
library(patchwork)
library(cowplot)

ROI <- readRDS("D:\\calc_temp\\field_test\\HGSC2_ROI_SO.rds")
ROI_spWAVE <- readRDS("D:\\calc_temp\\field_test\\HGSC2_ROI_spWAVE_c2c.rds")


L12R4_field <- perform_field_extract(ROI_spWAVE,"CXCL12.CXCR4")
L12R4_field %<>% mutate(subtype=ROI@meta.data$subtype)

p1 <- 
plot_field_direction2(L12R4_field,point="KU",
  hexified=T,hex_binwidth=c(20,20),
  arrow_color="black",
  arrow_linewidth = 0.5,
  #arrow_sf=0.001
  )+
  scale_fill_gradient2(low="#3B50C5", mid="#DCDCDC",high="#AD0C27")
p2 <- 
plot_field_direction2(L12R4_field,
  arrow_color="black",
  arrow_linewidth = 0.5,
  point="subtype",point_size=0.5)

L16R6_field <- perform_field_extract(ROI_spWAVE,"CXCL16.CXCR6")
L16R6_field %<>% mutate(subtype=ROI@meta.data$subtype)

p3 <- 
plot_field_direction2(L16R6_field,point="KU",
  hexified=T,hex_binwidth=c(20,20),
  arrow_color="black",
  arrow_linewidth = 0.5,
  #arrow_sf=0.001
  )+
  scale_fill_gradient2(low="#3B50C5", mid="#DCDCDC",high="#AD0C27",midpoint=0.4)

p4 <- 
plot_field_direction2(L16R6_field,
  arrow_color="black",
  arrow_linewidth = 0.5,
  point="subtype",point_size=0.5)

p_all <- (p1|p2)/(p3|p4)

final_plot <- ggdraw(p_all) +
  draw_label("Field Potential", x = 0.23, y = 0.98, size = 12, fontface = "bold") +
  draw_label("Subtype", x = 0.73, y = 0.98, size = 12, fontface = "bold") +
  draw_label("CXCL12-CXCR4", x = 0.005, y = 0.77, angle = 90, size = 12, fontface = "bold") +
  draw_label("CXCL16-CXCR6", x = 0.005, y = 0.27, angle = 90, size = 12, fontface = "bold")
final_plot

```

Since we directly utilized the 'point' parameter in the 'plot_field_direction2' function to visualize cell types, 
the resulting background differs from the first example, presenting a white background instead. 
In contrast, the black background is generated by the plotting function from the Seurat package.

Notably, the point parameter can also accept discrete variables such as character vectors and factors. 
And the value should be a column in the field result data.frame, like the 'subtype' in our example.
In such cases, different colors can be assigned by providing a named vector to the 'point_color' parameter.

