---
title: "Field_Est"
output: rmarkdown::html_vignette
mainfont: Arial
vignette: >
  %\VignetteIndexEntry{Field_Est}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  comment = "#>",
  root.dir = './'
)
```
This vignette outlines the steps of  of cell-cell communication network for **a single dataset using CellChat**. We showcase CellChat’s diverse functionalities by applying it to a scRNA-seq data on cells from lesional (LS, diseased) human skin from patients.  

CellChat requires gene expression data of cells as the user input and models the probability of cell-cell communication by integrating gene expression with prior knowledge of the interactions between signaling ligands, receptors and their cofactors. 

Upon infering the intercellular communication network, CellChat provides functionality for further data exploration, analysis, and visualization. 


## Load the required libraries
```{r setup}
library(magrittr)
library(dplyr)
library(tidyr)
library(stringr)

library(Seurat)
library(spWAVE)
```
Personally, I prefer to directly use 'library(tidyverse)' to load the tidyverse packages.
However, is not 